---
title: 'How to: Create folders'
author: "Jonathan Tan"
date: "08/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(
  position = c('top', 'right'),
  color = "auto",
  tooltip_message = "Copy code",
  tooltip_success = "Copied!"
)
```

<br>

## Starting a New Project

```{r, eval = FALSE}
Click: Create a project
Click: New Directory
Click: New Project
Click: New R Script
```

<br>

## Organising your work with folders!

Folders can be created to keep the main Rproject working directory organised and tidy:

```{r, eval = FALSE}
setwd("/run/media/jon/FILES/Work/Desktop items/Github_repos/")
dir.create("New_project")
setwd("/run/media/jon/FILES/Work/Desktop items/Github_repos/New_project")
dir.create("scripts")
dir.create("plots")
dir.create("data")
dir.create("markdown")
```

Under this folder structure:

+ A markdown script would be located in the "markdown" folder
+ A data file (eg. .csv, .pzfx) would be located in the "data" folder
+ Relative to the the markdown script, a file pathway to the .csv file would be:
../data/example.csv

***
