---
title: 'How to: Import a .csv file'
author: "Jonathan Tan"
date: "08/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(
  position = c('top', 'right'),
  color = "auto",
  tooltip_message = "Copy code",
  tooltip_success = "Copied!"
)
```

## Organising your work

Saving the .csv file in the working Rproject folder is a good strategy for keeping data organised and the workspace tidy.

Folders can be created using the following commands:
```{r, eval = FALSE}
setwd("/run/media/jon/FILES/Work/Desktop items/Github_repos/")
dir.create("New_project")
setwd("/run/media/jon/FILES/Work/Desktop items/Github_repos/New_project")
dir.create("scripts")
dir.create("plots")
dir.create("data")
dir.create("markdown")
```

Under this folder structure, the markdown script is located in the "markdown" folder. The .csv file is located in the "data" folder.
Therefore, relative to the the markdown script, the file pathway to the .csv file is:
../data/NHMRC_scoring_history2.csv

## Configure the .csv data

Using the "Files" navigator panel, open the "data" folder and click in the .csv file. Choose "Import dataset..."


Under "Data preview", the type of data can be modified:

+ Nominal data (like names) = character (text)
+ Height (tall/short) would be categorical data, but there's an order to it (ordinal) = factor
+ Age (whole number, integer)
+ Weight (any number even between whole numbers(integers), numeric/continuous)
+ Logical: true or false

Under "Import options":

+ Rename dataframe (nhmrc)
+ Select any other relevant options

Under "Code preview":

+ Copy code. This will be pasted into the script (below)

## Import the .csv data

The code which has been copied is then used to import the .csv file into a dataframe called "nhmrc":
```{r}
library(readr)
nhmrc <- read_csv("../data/NHMRC_scoring_history2.csv", 
                  col_types = cols(Score = col_number()))
head(nhmrc, n = 12L)
```

