---
title: 'How to: Enter data into RStudio'
author: "Jonathan Tan"
date: "12/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(
  position = c('top', 'right'),
  color = "auto",
  tooltip_message = "Copy code",
  tooltip_success = "Copied!"
)
```


<br>

## Understanding the data

Example: Compare the percentage of myeloid-biased HSC between young and old animals.

<br>

When thinking about your data, identify the following:

1. What are you comparing **(independent variables)**?

- Young vs old animals (ie. Category = Group)

2. What is your measurement **(dependent variables)**?
  * Percent myeloid-biased HSC (Numerical value)

<br>
  
## Create a datafile directly in RStudio

http://www.sthda.com/english/articles/24-ggpubr-publication-ready-plots/76-add-p-values-and-significance-levels-to-ggplots/#compare-two-paired-samples

http://www.sthda.com/english/wiki/paired-samples-t-test-in-r



Data can be entered into two numeric vectors:
```{r}
young <-c(30, 38.9, 31.7, 2.38, 8.11, 33.7, 28.7)
old <-c(50.7, 78.6, 60, 36.9, 24.3, 36.9, 58.2)
```

<br>

Create a data frame called "expt":
```{r}
expt <- data.frame( 
  group = rep(c("young", "old"), each = 7),
  HSC = c(young,  old)
)

print(expt)
```

<br>

***
