---
title: 'How to: Perform a one-sample t-test'
author: "Jonathan Tan"
date: "18/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(
  position = c('top', 'right'),
  color = "auto",
  tooltip_message = "Copy code",
  tooltip_success = "Copied!"
)
```

<br>

http://www.sthda.com/english/wiki/one-sample-t-test-in-r


<br>

## One-sample t-test

Used to compare the **mean** of one sample to a **known standard**

+ (or a theoretical or hypothetical mean)

<br>

eg. For an experiment where you have *control* and *treatment* conditions.

+ If data is expressed as "percent of control", you can test whether the average value of the treatment condition differs significantly from 100.

<br>

*Note that a one-sample t-test can be used only when the data are normally distributed. This can be checked using the Shapiro-Wilk test.*

<br>
  
## Create a datafile directly in RStudio

In this example, we are testing whether a 10min incubation is different to a 30min incubation.

+ The 10min incubation is the control and is given a standard value of 100% (mu).
+ The treatment in question is a 30min incubation.

<br>

Data can be entered into a numeric vector:
```{r}
data <-c(89, 102.3, 125)
```

<br>

A data frame is then created called "expt":
```{r}
expt <- data.frame( 
  percent = data
)

print(expt)
```

<br>

## Statistics

Does the data follow a normal distribution? Use the Shapiro-Wilk normality test:

+ Null hypothesis: the data are normally distributed
+ Alternative hypothesis: the data are not normally distributed

```{r}
shapiro.test(expt$percent)
```
From the output, the p-value is greater than the significance level 0.05 implying that the distribution of the data are not significantly different from normal distribution. In other words, we can assume normality.

<br>

## Compute one-sample t-test:
The specified value for the known standard (mu) is 100%.
```{r}
test <- t.test(expt$percent, mu = 100)
test
```
<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 10px;}
</style>
<div class = "blue">

The p-value of the test is **0.6567**, which is greater than the significance level **alpha = 0.05**.

We can conclude that the mean percentage of the 30min incubation is not significantly different to 10mins.

</div>


<br>

***




