---
title: 'How to: Generate Descriptive Statistics'
author: "Jonathan Tan"
date: "08/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(
  position = c('top', 'right'),
  color = "auto",
  tooltip_message = "Copy code",
  tooltip_success = "Copied!"
)
```

## Understanding your data

Example: Comparing CD150hi (myeloid-biased) HSC between young and old animals

When thinking about your data, identify the following:

1. What are you comparing **(independent variables)**?
 + *Young and old animals (ie. Category = Age)*

2. What is your measurement **(dependent variables)**?
  + *The percentage of CD150^hi^ HSC (Category = HSC)*
  
## Import data from a Prism file
List all the data tables in the Prism file.

```{r}
library(pzfx)
pzfx_tables("../data/young vs old graphs JTmod.pzfx")
```

Now we can select which data table to import:
```{r}
Fig423b <- read_pzfx("../data/young vs old graphs JTmod.pzfx", "Figure 4.2.3B")
head(Fig423b, N = 6)
```

Oh no! The data is in a format which RStudio can't read!

We need to arrange this dataset so that each column lists a variable (Age, HSC).

Fortunately, we can call on the "pivot_longer" function to convert data into "Tall format":
```{r, message = FALSE}
library(tidyverse)
Fig423b.tall <- Fig423b %>% select(Young, Old, "Old asplenic") %>%
  pivot_longer(., cols = c(Young,Old,"Old asplenic"), names_to = "Age", values_to = "HSC") ## Concatenate columns (Young, Old, Old Asplenic) into one column (age)
head(Fig423b.tall, N = 21)
```

Now all the observations (values for percent HSC) are listed in one column.

## Run descriptive statistics
We will now find the average %HSC for all treatment groups (Average = mean(HSC)).
```{r}
Fig423b.tall %>% 
  select(Age, HSC) %>%
  filter(Age == "Young"|
           Age == "Old"|
           Age == "Old asplenic") %>% 
  na.omit() %>% # Omit values with 'NA'
  group_by(Age) %>% 
  summarise(Average = mean(HSC))
```

